---
title: "A first look at the Divvy data"
author: "Peter Carbonetto"
date: 2017-06-29
output: html_document
---

<!-- Define defaults shared by all workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository
 exists and R package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

I begin by loading a few packages, as well as some additional
functions I wrote for this project, into the R environment.

```{r load-packages, message = FALSE}
library(data.table)
source("../code/functions.R")
```

<br>

## Reading the data

I wrote a function, `read.divvy.data`, that reads in the trip and
station data from the CSV files that were downloaded from the Divvy
website. This function uses `fread` from the `data.table` package,
which is much faster than the `read.table` function for reading CSV
files. This function takes a few additional steps to prepare the data
so that they are in a format that is easier to work with.

```{r load-data}
divvy <- read.divvy.data()
```
<br>

## A first glance at the Divvy data

```{r stations-overview}
nrow(divvy$stations)
ncol(divvy$stations)
names(divvy$stations)
```

```{r trips-overview}
nrow(divvy$trips)
ncol(divvy$trips)
names(divvy$trips)
```

- Number of stations
- Number of trips in 2016

Which station(s) had the most activity?

```{r most-trips-by-station}
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
