---
title: "A first glance at the Divvy data"
author: "Peter Carbonetto"
date: 2017-06-29
output: html_document
---

<!-- Define defaults shared by all workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository
 exists and R package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

Here, we will take a brief look at the data provided by Divvy.

I begin by loading a few packages, as well as some additional
functions I wrote for this project.

```{r load-packages, message = FALSE}
library(data.table)
source("../code/functions.R")
```
<br>

## Reading the data

I wrote a function, `read.divvy.data`, that reads in the trip and
station data from the
[Divvy CSV files](https://www.divvybikes.com/system-data). This
function uses `fread` from the `data.table` package to quickly read in
the data (it is much faster than `read.table`). This function also
prepares the data, notably the dates and times, so that they are
easier to work with.

```{r load-data}
divvy <- read.divvy.data()
```
<br>

## A first glance at the Divvy data

We have information on 581 Divvy stations across the city.

```{r stations-overview}
nrow(divvy$stations)
print(head(divvy$stations),row.names = FALSE)
```

We also have information about the >3 million trips taken on Divvy
bikes in 2016.

```{r trips-overview}
nrow(divvy$trips)
print(head(divvy$trips),row.names = FALSE)
```

Out of all the Divvy stations in Chicago, the one on Navy Pier (near
the corner of Streeter and Grand) had the most activity by far.

```{r most-trips-by-station}
departures <- table(divvy$trips$from_station_name)
as.matrix(head(sort(departures,decreasing = TRUE)))
```
<br>

## Divvy bikes at the University of Chicago

I will also take a close look at trip data for the main Divvy station
on the University of Chicago campus, since this is where I work.

```{r uchicago-station}
sum(divvy$trips$from_station_name == "University Ave & 57th St",na.rm = TRUE)
```
<br>

## Session information

This is the version of R and the packages that were used to generate
these results.

```{r session-info}
```
