---
title: "A first look at the Divvy data"
author: "Peter Carbonetto"
date: 2017-06-29
output: html_document
---

<!-- Define defaults shared by all workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository
 exists and R package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

Begin by loading packages and some useful function definitions into
the R environment.

```{r load-packages, message = FALSE}
library(data.table)
source("../code/functions.R")
```

## Read the data

I wrote a function, `read.divvy.data` to read the trip and station
data from the CSV files. It uses the `fread` function from the
`data.table` package which is much faster than the `read.table`
function. It also does a few additional data processing steps so that
the data are in a format that are easier to work with for our
analyses.

```{r load-data}
divvy <- read.divvy.data()
```
<br>

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
