---
title: "Divvy usage by time-of-day"
author: "Peter Carbonetto"
output: html_document
---

<!-- Define defaults shared by all workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository
 exists and R package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

Here we use the Divvy trip data to examine biking trends over the
course of the day.

I begin by loading a few packages, as well as some additional
functions I wrote for this project.

```{r load-packages, message = FALSE}
library(data.table)
library(ggplot2)
source("../code/functions.R")
```
<br>

## Read the data

Following [my earlier steps](first-glance.html), I use function
`read.divvy.data` to read the trip and station data from the CSV
files.

```{r load-data, cache = TRUE}
divvy <- read.divvy.data()
```

To make it easier to compile statistics by time-of-day, I convert the
"start hour" column to a factor (categorical variable).

```{r convert-start-hour}
divvy$trips <- transform(divvy$trips,
                         start.hour = factor(start.hour,0:23))
```
<br>

## Count departures

Add text here.

```{r timeofday-barchart}
ggplot(divvy$trips,aes(start.hour)) +
  geom_bar(fill = "black",width = 0.6) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

Add text here.

```{r timeofday-barchart-2}
ggplot(divvy$trips,aes(start.hour)) +
  geom_bar(fill = "black",width = 0.6) +
  facet_wrap(~start.day,ncol = 2) +
  scale_x_discrete(breaks = seq(0,24,2)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
<br>

## University of Chicago trends

```{r timeofday-uchicago-barchart}
ggplot(subset(divvy$trips,from_station_name == "University Ave & 57th St"),
       aes(start.hour)) +
  geom_bar(fill = "black",width = 0.6) +
  facet_wrap(~start.day,ncol = 2) +
  scale_x_discrete(breaks = seq(0,24,2)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

<br>

## Session information

This is the version of R and the packages that were used to generate
these results.

```{r session-info}
```
