---
title: "Divvy usage by time-of-day"
author: "Peter Carbonetto"
output: html_document
---

<!-- Define defaults shared by all workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository
 exists and R package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

Here we use the Divvy trip data to examine biking trends over the
course of a typical day in Chicago. 

I begin by loading a few packages, as well as some additional
functions I wrote for this project.

```{r load-packages, message = FALSE}
library(data.table)
library(ggplot2)
source("../code/functions.R")
```

## Read the data

Following [my earlier steps](first-glance.html), I use function
`read.divvy.data` to read the trip and station data from the CSV
files.

```{r load-data}
divvy <- read.divvy.data()
```

To make it easier to compile statistics by time of day, I convert the
"start hour" column to a factor.

```{r convert-start-hour}
divvy$trips <- transform(divvy$trips,start.hour = factor(start.hour,0:23))
```

## Count departures by time-of-day

Now that `start.hour` is a factor, it is easy to create a bar chart
showing the total number of departures at each hour. Unsurprisingly,
we see little biking activity at night. Further, the two peaks
("modes") in the bar chart nicely recapitulate the morning and
afternoon rush hours.

```{r timeofday-barchart, fig.height = 2.5, fig.width = 5}
ggplot(divvy$trips,aes(start.hour)) +
  geom_bar(fill = "dodgerblue",width = 0.6) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

However, this bar chart is a bit muddled because it is counting trips
during the week and on the weekends. Consider that the bin count
$x[h]$ for hour $h$ in the histogram above is a sum of the counts for
each day of the week:

$$
\begin{align}
x[h] 
& = \sum_{i\;\in\;\mathsf{DaysOfTheWeek}} x_i[h] \\
& = x_{\mathsf{Mon}}[h] + x_{\mathsf{Tue}}[h] + x_{\mathsf{Wed}}[h] +
x_{\mathsf{Thu}}[h] + x_{\mathsf{Fri}}[h] + x_{\mathsf{Sat}}[h] +
x_{\mathsf{Sun}}[h]
\end{align}
$$

*Note:* The math above is embedded in the webpage using 
[MathJax](https://github.com/mathjax/MathJax). See [here](https://math.meta.stackexchange.com/questions/5020/mathjax-basic-tutorial-and-quick-reference)
for an excellent reference on MathJax.

Once we plot the counts separately for each the day of the week, the
rush-hour trends become more obvious. (Also notice that the rush-hour
weeks disappear on Saturday and Sunday.)

```{r timeofday-barchart-2, fig.height = 6, fig.width = 6}
ggplot(divvy$trips,aes(start.hour)) +
  geom_bar(fill = "dodgerblue",width = 0.75) +
  facet_wrap(~start.day,ncol = 2) +
  scale_x_discrete(breaks = seq(0,24,2)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

## Time-of-day trends at University of Chicago

The commuting trends are different at the University of Chicago Divvy
station---there isn't much of a morning rush hour. This may be because
students and staff don't regularly use the Divvy bikes for commuting.

```{r timeofday-uchicago-barchart, fig.height = 6, fig.width = 6}
ggplot(subset(divvy$trips,from_station_name == "University Ave & 57th St"),
       aes(start.hour)) +
  geom_bar(fill = "dodgerblue",width = 0.75) +
  facet_wrap(~start.day,ncol = 2) +
  scale_x_discrete(breaks = seq(0,24,2)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

## Session information

This is the version of R and the packages that were used to generate
these results.

```{r session-info}
```
